---
title: "Problem Set1"
author: "Himani Anil Deshpande"
date: "9/7/2021"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lattice)
library(readxl)
```


## Loading the first four columns from the Dataset

```{r}

data <- read_excel("Assignment 1/DrivingdataAll.xls", range = cell_cols("A:D"))


#View(data)
```



## Analysing the different Odometer readings using Histograms

## Baseline Odometer Reading
```{r}
ggplot(data, aes(x = `Odom Reading 1 (Previous)`)) + 
  geom_histogram(fill = 'green') +
  xlab("Baseline odometer reading (in miles)") +
  ylab("Count of Records") +
 ggtitle("Histogram of baseline odometer reading") +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

## Self-reported Odometer Reading
```{r}
ggplot(data, aes(x = `Odom Reading 1 (Update)`)) + 
  geom_histogram(fill = 'blue') +
  xlab("Self-reported odometer reading (in miles)") +
  ylab("Count of Records") +
  ggtitle("Histogram of self-reported odometer reading") +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

Both the histograms are right skewed. So, the mean is greater than median. 

## Performing Log transformation to further anlayze the distribution

## Baseline Odometer Reading
```{r}
ggplot(data, aes(x = `Odom Reading 1 (Previous)`)) + 
  geom_histogram(fill = 'green') + 
  scale_x_log10() +
  xlab("Log of Baseline odometer reading") +
  ylab("Count of Records") +
 ggtitle("Histogram of baseline odometer reading") + 
  labs(subtitle = "Log Transformation of Baseline Odometer") +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```




## Self-reported Odometer Reading
```{r}
ggplot(data, aes(x = `Odom Reading 1 (Update)`)) + 
  geom_histogram(fill = 'blue') + scale_x_log10() +
  xlab("Log of Self-reported odometer reading") +
  ylab("Count of Records") +
  ggtitle("Histogram of self-reported odometer reading") + 
  labs(subtitle = "Log Transformation of Self-Reported Odometer") +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

The log transformation makes the data left skewed. So we cannot use mean for analysis.





 As mentioned in the paper, The  response  variable  was  the  difference  between  self-reported  miles  and  baseline  miles(Column D minus Column C.)  So I am going to analyze the response  variable and its histogram


## Analyzing the reponse Variable

```{r}

ggplot(data,  aes(y =`Odom Reading 1 (Update)`- `Odom Reading 1 (Previous)` ,
                  x = `OMR Version`, col = `OMR Version`)) + 
  ylab("Response variable") +
  ggtitle("Scatter Plot of Response V/S OMR Version") +
             geom_jitter(position = position_jitter(height = 0, width=0.4))


```

The batches share the same range
So I will check the spread using the histogram



## Histogram of Response Variable

```{r}
ggplot(data, aes(x =`Odom Reading 1 (Update)`- `Odom Reading 1 (Previous)`)) +      
  geom_histogram(fill = 'black') + 
  xlab("Response variable (in miles)") +
  ylab("Count of Records") +
 ggtitle("Histogram of Response variable ") +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

## Trying Histogram with different Bin sizes

```{r}
ggplot(data, aes(x =`Odom Reading 1 (Update)`- `Odom Reading 1 (Previous)`)) +      
  geom_histogram(fill = 'black', breaks = seq(-1, 50500, 1000)) +
  xlab("Response variable (in miles)") +
  ylab("Count of Records") +
 ggtitle("Histogram of Response variable ") +
   theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```


## Histogram of Reponse variable categorized over OMR Version

```{r}

ggplot(data, aes(x = `Odom Reading 1 (Update)`- `Odom Reading 1 (Previous)`)) + 
  geom_histogram() +
  xlab("Response variable (in miles)") +
  ylab("Count of Records") +
 ggtitle("Histogram of Response variable ") +
  labs(subtitle = "Categorized on OMR Version") +
  facet_grid(~`OMR Version`)
```

The histogram is not skewed and look symmetric So we dont need to perform any transformations for understanding the distribution of response variable.
The histogram of response variable( self reported - baseline odometer readings) clearly shows that it is a uniform distribution.

So we will go ahead and check if the set of data comes from same distribution by plotting QQ plots.


## QQ Plots

## QQ Plot of Response variable

```{r}
ggplot(data, aes(sample = `Odom Reading 1 (Update)`- `Odom Reading 1 (Previous)`))+
  stat_qq(distribution = qunif) +
  ylab("Response Variable") +
  ggtitle("QQ Plot of Response variable") +
 labs(subtitle = "Uniform Distribution") 

```



## QQ plot of Response variable categorized over OMR Version


```{r}

ggplot(data, aes(sample = `Odom Reading 1 (Update)`- `Odom Reading 1 (Previous)`))+
  stat_qq(distribution = qunif) +
  ylab("Response Variable") +
  ggtitle("QQ Plot of Response variable over OMR version") +
  labs(subtitle = "Uniform Distribution") +
  facet_grid(~`OMR Version`) 




```




A straight line suggests a probability model, which implies that both the Self reported and Baseline Odometer reading came from the same distribution.
The sample shows an exact straight line for Uniform distribution, which is rare in a real world data. The real world data usually poses some outliers or skewness in data, however the dataset  we are working on doesnt have any outliers or any skewness with repect to response variable. All the data points falling on straight line with no deviation or curve implies that the data is fradulent.
It seems that the data was created specifically to fit a Uniform distribution.
So, any claim that the hypothesis is correct based on this data cannot be deemed trustworthy.



